---
title: "Clinic data"
author: "Lluís Revilla Sancho"
date: "`r date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, fig.width = 120, fig.height = 50, collapse = TRUE)
library("RGCCA")
library("gliomaData")
library("ggplot2")
library("patchwork")
library("integration")
theme_set(theme_bw())
```

# Introduction

In the other optimizations we used the same data as used in the RGCCA vignette. 
However further phenotype data is available as part of the clinic data from the 
gliomaData package, here we explore furhter if using these data improve the 
relation finding.

```{r introduction}
data(ge_cgh_locIGR, package = "gliomaData")
A <- ge_cgh_locIGR$multiblocks
Loc <- factor(ge_cgh_locIGR$y)
levels(Loc) <- colnames(ge_cgh_locIGR$multiblocks$y)
ncomp <- rep(1, length(A))
# rgcca algorithm using the dual formulation for X1 and X2 
# and the dual formulation for X3
A[[3]] = A[[3]][, -3]
```

# Phenotype

That is the default data used, now  we can use other data relevant to the 
disease:

```{r clinic}
data("clinic", package = "gliomaData")
# Remove any column with an NA or empty fields
keep <- vapply(clinic, function(x){!any(is.na(x) | x == "")}, logical(1L))
clinic <- clinic[, keep]
# Remove any value that is unique in all column or each row
keep <- vapply(clinic, function(x){
  y <- length(unique(x))
   y > 1 & y != length(x)}, logical(1L))
clinic <- clinic[, keep]
colnames(clinic)

# We can use the Sex, the age_an, the localization and grade.OMS
fact2numeric <- function(x){
  as.numeric(levels(x))[x]
}

clinic$sexe <- ifelse(clinic$sexe == "M", 1, 0)
clinic$age_an <- as.numeric(sub(",", ".", as.character(clinic$age_an), fixed = TRUE))
# An outlier of age that we assume that that those are days and not years, 
# so it is divided by 365
clinic$age_an[clinic$age_an > 100] <- clinic$age_an[clinic$age_an > 100]/365.25
Localisation <- sapply(levels(clinic$localisation), function(x){
  as.numeric(x == clinic$localisation)})
```
Now that we have ready our data we add it. Note that we add another change, instead of one numeric value for each localization, we have one column for each and 0 or 1 if present (but we remove one to avoid redundance).
```{r data}
A$y <- as.matrix(clinic[, c("sexe", "age_an", "grade.OMS")])
A$Loc <- as.matrix(Localisation[, 1:2]) # We remove the last column because it is redudant
names(A) <- c("GE", "CGH", "Invariable", "Localisation")
```

```{r}
library("globaltest")
library("SummarizedExperiment")
assays <- SimpleList("GE" = base::t(A$GE), "CGH" = base::t(A$CGH))
assays <- S4Vectors::sapply(assays, function(x){
  base::colnames(x) <- NULL
  x
})
colData <- DataFrame(clinic[, c("localisation", "grade.OMS", "age_an", "sexe")])
base::rownames(colData) <- NULL
se <- SummarizedExperiment(assays= assays, colData=colData)
```

# Srhinkage

```{r tau}
(min_shrinkage <- sapply(A, function(x) {
  1 / sqrt(ncol(x))
}))
shrinkage <- c(.071,.2, 1, 1)
```

# Design matrix

Now that we are ready to go, we need to test, all the previous things together.
Here we create a design matrix of dimension 4 with 4 weights.

```{r}
C_list2 <- weight_design(4, 4)
```

# Evaluation of designs

```{r}
testing <- function(x, type) {
  result.sgcca <- sgcca(A, x, c1 = shrinkage, ncomp = rep(1, length(A)), 
                       scheme = type, verbose = FALSE)
  
  ind <- apply(which(upper.tri(result.sgcca$C), arr.ind = TRUE), 1, 
               paste0, collapse = "")
  
  # Correlation between Y
  Y <- simplify2array(result.sgcca$Y, higher = FALSE)
  cY <- cor(Y)
  d <- cY * result.sgcca$C
  var <- cY[upper.tri(cY)]
  names(var) <- paste0("vs", ind)
  # Canonical correlation
  cc <- sum(d[upper.tri(d)])
  # Values of the design matrix
  vars <- result.sgcca$C[upper.tri(result.sgcca$C)]
  names(vars) <- paste0("var", ind)
  # weights used
  weight <- sum(vars != 0)
  names(weight) <- "weights"
  
  # Output
  c(var, unlist(result.sgcca$AVE[c("AVE_inner", "AVE_outer")]), cc1 = cc, 
    vars, weight)
}
```


```{r}
out <- sapply(C_list2, testing, type = "centroid")
out2 <- t(out)
def <- as.data.frame(out2)
def$weights <- as.factor(def$weights)
offset <- is.na(def$AVE_inner)
def <- droplevels(def[!offset, ])
rownames(def) <- seq_len(nrow(def))
saveRDS(def, file = "centroid_complicated.RDS")
```

# Exploring results

We can explore by type of model how accurate is the model.
```{r}
library("dplyr")
def %>% 
  group_by(weights) %>% 
  add_count() %>% 
  dplyr::filter(AVE_inner == max(AVE_inner)) %>% 
  arrange(desc(as.numeric(weights)), desc(AVE_inner)) %>% 
  select(-matches("vs"))
```
We can see that the model with just 2 weights is the most accurate to explain the model.
Which is the one where gene expression is related to localization and CGH is related to the demographics.
We see that the most complex model doesn't mean more explanation of the variables.
We can look at what happens at the extremes, when the interaction is zero:
```{r}
o <- sapply(def[, 10:15], function(x){
  y <- droplevels(def[x == 0, ])
  y$weights <- as.numeric(as.character(y$weights))
  y[y$AVE_inner == max(y$AVE_inner), , drop = TRUE]
})
o2 <- apply(o, 1, unlist)
as_tibble(unique(o2))
```

We can look at what happens at the extremes, when the interaction is 1:
```{r}
o <- sapply(def[, 10:15], function(x){
  y <- droplevels(def[x == 1, ])
  y$weights <- as.numeric(as.character(y$weights))
  y[y$AVE_inner == max(y$AVE_inner), , drop = TRUE]
})
o2 <- apply(o, 1, unlist)
as_tibble(unique(o2))
```

We can look further to the graphs:
```{r}
def2 <- def %>% mutate(weight = (var12+var13+var14+var23+var24+var34))
p <- ggplot(def2)
r1 <- p + geom_point(aes(cc1, AVE_inner, col = weights)) + guides(col = FALSE)
r2 <- p + geom_point(aes(cc1, AVE_outer, col = weights)) + guides(col = FALSE)
r3 <- p + geom_point(aes(AVE_inner, AVE_outer, col = weights))
p + geom_point(aes(AVE_inner, AVE_outer, col = weights)) + facet_wrap(~weights)
per12 <- p + 
  geom_point(aes(AVE_inner, var12/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var12")
per13 <- p + 
  geom_point(aes(AVE_inner, var13/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var13")
per14 <- p + 
  geom_point(aes(AVE_inner, var14/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var14")
per23 <- p + 
  geom_point(aes(AVE_inner, var23/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var23")
per24 <- p + 
  geom_point(aes(AVE_inner, var24/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var24")
per34 <- p + 
  geom_point(aes(AVE_inner, var34/weight, col = weights)) + 
  ylim(c(0, 1)) + ylab("% var34")
q1 <- p + geom_point(aes(var12, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var12, AVE_outer, col = weights))
q2 <- p + geom_point(aes(var13, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var13, AVE_outer, col = weights))
q3 <- p + geom_point(aes(var14, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var14, AVE_outer, col = weights))
q4 <- p + geom_point(aes(var23, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var23, AVE_outer, col = weights))
q5 <- p + geom_point(aes(var24, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var24, AVE_outer, col = weights))
q6 <- p + geom_point(aes(var34, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var34, AVE_outer, col = weights))

s1 <- p + geom_point(aes(var1, AVE_inner, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var1, AVE_inner, col = weights))
s2 <- p + geom_point(aes(var2, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_inner, col = weights))
s3 <- p + geom_point(aes(var3, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_inner, col = weights))
(r1 + r2 + r3) / (q1 + q2 + q3) / (s1 + s2 + s3)
```

# Testing

```{r}
design.final <- matrix(0, nrow = length(A), ncol = length(A), 
                       dimnames = list(names(A), names(A)))
design.final[1, 4] <- design.final[4, 1] <- 1
design.final[2, 3] <- design.final[3, 2] <- 1/3
result.sgcca <- sgcca(A, design.final, c1 = shrinkage, ncomp = rep(2, length(A)), 
                       scheme = "centroid", verbose = FALSE)
plot(result.sgcca$Y[[1]][, 1], result.sgcca$Y[[2]][, 1], xlab = "GE", ylab = "CGH", 
     col = clinic$localisation)

```


```{r}
lmC <- lm(AVE_inner~0+var12+var13+var14+var23+var24+var34, data = def)
glance(lmC)
tidy(lmC) %>% arrange(desc(estimate))
```


# SessionInfo {-}

```{r, echo=FALSE}
devtools::session_info()
```

