---
author: "Lluís Revilla Sancho"
title: "SGCCA hyperparameters"
date: "`r date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, fig.width = 120, fig.height = 50, collapse = TRUE)
library("RGCCA")
```

# Introduction

This project page is to analyze the effect of some parameters on the sparse
generalized canonical correlation analysis (SGCCA) from the [RGCCA](https://cran.r-project.org/package=RGCCA) package^[It doesn't use my slightly modified [version](https://github.com/llrs/RGCCA)].


# Sections

The arguments needed for a sgcca are:
```{r}
args(sgcca)
```

 - A
 Is the input of the several blocks of variables to be related.
 - **C**  
 Is the design of the model, with more data more models can be used. For each model several weights can be used too. This is explored [here](weight.html)
 - **c1**  
 The regularizer parameter, also called tau in the RGCCA method. Its effect is explored [here](tau.html)
 - **scheme**  
 How are the several canonical correlations added up in the optimization function:
    + horst  
    The canonical correlation as is.
    + centroid  
    The absolute canonical correlation
    + factorial  
    The square of the canonical correlation
    + free  
    One can input any function
 - scale  
 Parameter to standarize to zero means and unit variance and divided by the square root of its number of variables.
 - init  
 Starting point for the optimization function, an be both `sva` and `random`
 - bias  
 Use biaised or unbiaised estimator of the variance and covariance
 - tol  
 Precision of the numeric approximation.
 - verbose  
 Prints messages along the calculation and plots the optimization approach.
 
