---
title: "Weight optimization"
author: "Lluís Revilla Sancho"
date: "`r date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, fig.width = 120, fig.height = 50, collapse = TRUE)
library("RGCCA")
library("gliomaData")
library("ggplot2")
library("patchwork")
library("integration")
theme_set(theme_bw())
```

# Introduction

In this vignette we explore the relationship between weights (and models) with the outcome of the integration using the glioma data available.

```{r introduction}
data(ge_cgh_locIGR, package = "gliomaData")
A <- ge_cgh_locIGR$multiblocks
Loc <- factor(ge_cgh_locIGR$y)
levels(Loc) <- colnames(ge_cgh_locIGR$multiblocks$y)
ncomp <- rep(1, length(A))
shrinkage <- c(.071,.2, 1)
# rgcca algorithm using the dual formulation for X1 and X2 
# and the dual formulation for X3
A[[3]] = A[[3]][, -3]

# Prepare some preliminary design/weights
C <-  matrix(c(0, 0, 1, 0, 0, 1, 1, 1, 0), 3, 3)
dimnames(C) <- list(names(A), names(A))
```

# Weights

To test some weights we can use the RV index between the matrices

```{r RV, echo=FALSE}
D2 <- D <- C
for (i in seq_len(nrow(C))) {
  Ai <- scale(A[[i]], scale = FALSE)
  for (j in seq_len(i)) {
    Bj <- scale(A[[j]], scale = FALSE)
    if (i == j) {
      D2[i, i] <- 0
    } else {
      D2[i, j] <- MatrixCorrelation::RV(Ai, Bj)
      D2[j, i] <- D2[i, j]
    }
  }
}

```

The correlation between the first components of a multidimentional scaling plot:

```{r cor}
cmds <- lapply(A, function(x){cmdscale(dist(scale2(x)), k = 1)})
for (i in seq_len(nrow(C))) {
  for (j in seq_len(i)) {
    if (i == j) {
      D[i, i] <- 0
    } else {

      D[i, j] <- cor(cmds[[i]], cmds[[j]])
      D[j, i] <- D[i, j]
    }
  }
}
D <- abs(D)
```


Create the 1000 design matrix combinations

```{r}
nweight <- 11
weight <- seq(from = 0, to = 1, length.out = nweight)
C_list2 <- weight_design(11, 3)
keep <- check_design(C_list2)
C_list2 <- C_list2[keep]
```


```{r append}
C_list2 <- c(C_list2, list(D, D2))
```

## The genes selected

To know if it really makes sense to change the matrix weights we can look for changes

```{r}
result.sgcca = sgcca(A, D, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
result.sgcca2 = sgcca(A, D2, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
result.sgcca3 = sgcca(A, C, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
comp.weight <- function(x, y){
  xdiff <- x != 0
  ydiff <- y != 0
  intersect(names(x[xdiff]), names(y[ydiff]))
}

l <- list(result.sgcca, result.sgcca2, result.sgcca3)
outm <- matrix(0, ncol = length(l), nrow = length(l))
for (i in seq_len(length(l))) {
  for (j in seq_len(i)) {
    outm[i, j] <- length(comp.weight(l[[i]]$a[[1]][, 1], l[[j]]$a[[1]][, 1]))
    outm[j, i] <- outm[i, j]
  }
}
GE <- outm/max(outm)
dimnames(GE) <- list(c("D", "D2", "C"), c("D", "D2", "C"))
GE
```
We can see that with different weights different genes are selected, and if we look at the CGH:
```{r}
outm <- matrix(0, ncol = length(l), nrow = length(l))
for (i in seq_len(length(l))) {
  for (j in seq_len(i)) {
    outm[i, j] <- length(comp.weight(l[[i]]$a[[2]][, 1], l[[j]]$a[[2]][, 1]))
    outm[j, i] <- outm[i, j]
  }
}
CGH <- outm/max(outm)
dimnames(CGH) <- list(c("D", "D2", "C"), c("D", "D2", "C"))
CGH
```

We see they also change however almost all agree but there are some slight changes of up to 25%


```{r}
testing <- function(x, type) {
  result.sgcca <- RGCCA::sgcca(A, x, c1 = shrinkage, ncomp = c(1, 1, 1), 
                       scheme = type, verbose = FALSE)
  
  analyze(result.sgcca)
}

# Design matrix:
# 0.0 var12 var13
# var12 0.0  var23
# var13 var23  0.0
```

### Centroid scheme

Using the centroid scheme

```{r centroid}
out <- sapply(C_list2, testing, type = "centroid", USE.NAMES = FALSE)
out2 <- t(out)
def <- as.data.frame(out2)
def$weights <- as.factor(def$weights)
offset <- is.na(def$AVE_inner)
centroid_weights <- droplevels(def[!offset, ])
saveRDS(centroid_weights, file = "centroid_weights.RDS")
```


```{r}
library("broom")
lmWC0 <- lm(AVE_inner ~ 0+var3+var2+var1, data=centroid_weights)
lmWC <- lm(AVE_inner ~ 0+var3*var2*var1, data=centroid_weights)
anova(lmWC, lmWC0)
glance(lmWC)
tidy(lmWC) %>% arrange(desc(estimate))
```

We can plot them

```{r plot_centroid}
p <- ggplot(centroid_weights)
r1 <- p + geom_point(aes(cc1, AVE_inner, col = weights)) + guides(col = FALSE)
r2 <- p + geom_point(aes(cc1, AVE_outer, col = weights)) + guides(col = FALSE)
r3 <- p + geom_count(aes(AVE_inner, AVE_outer, col = weights))
q1 <- p + geom_point(aes(var1, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q2 <- p + geom_point(aes(var2, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q3 <- p + geom_point(aes(var3, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_outer, col = weights))
s1 <- p + geom_point(aes(var1, AVE_inner, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var1, AVE_inner, col = weights))
s2 <- p + geom_point(aes(var2, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_inner, col = weights))
s3 <- p + geom_point(aes(var3, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_inner, col = weights))
(r1 + r2 + r3) / (q1 + q2 + q3) / (s1 + s2 + s3)
```

And we can see the stability for each type of model
```{r}
library("dplyr")
centroid_weights %>% 
  group_by(weights) %>% add_count() %>% 
  dplyr::filter(AVE_inner == max(AVE_inner)) %>% 
  arrange(desc(AVE_inner), as.numeric(weights)) %>% 
  select(-matches("vs"))
```

We can do the same with horst scheme.


### Horst scheme

Using horst scheme

```{r horst}
out <- sapply(C_list2, testing, type = "horst")
out2 <- t(out)
def <- as.data.frame(out2)
def$weights <- as.factor(def$weights)
offset <- is.na(def$AVE_inner)
horst_weights <- droplevels(def[!offset, ])
saveRDS(horst_weights, file = "horst_weights.RDS")
```


```{r}
lmWH0 <- lm(AVE_inner ~ 0+var3+var2+var1, data=horst_weights)
lmWH <- lm(AVE_inner ~ 0+var3*var2*var1, data=horst_weights)
anova(lmWH0, lmWH)
glance(lmWH)
tidy(lmWH) %>% arrange(desc(estimate))
```

```{r plot_horst}
p <- ggplot(horst_weights)
r1 <- p + geom_point(aes(cc1, AVE_inner, col = weights)) + guides(col = FALSE)
r2 <- p + geom_point(aes(cc1, AVE_outer, col = weights)) + guides(col = FALSE)
r3 <- p + geom_count(aes(AVE_inner, AVE_outer, col = weights))
q1 <- p + geom_point(aes(var1, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q2 <- p + geom_point(aes(var2, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q3 <- p + geom_point(aes(var3, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_outer, col = weights))
s1 <- p + geom_point(aes(var1, AVE_inner, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var1, AVE_inner, col = weights))
s2 <- p + geom_point(aes(var2, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_inner, col = weights))
s3 <- p + geom_point(aes(var3, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_inner, col = weights))
(r1 + r2 + r3) / (q1 + q2 + q3) / (s1 + s2 + s3)
```

```{r}
horst_weights %>% 
  group_by(sign(vs12), sign(vs13), sign(vs23), weights) %>% add_count() %>% 
  filter(AVE_inner == max(AVE_inner)) %>% 
  arrange(as.numeric(weights))
```

### Factorial scheme

Using the factorial scheme:

```{r factorial}
out <- sapply(C_list2, testing, type = "factorial")
out2 <- t(out)
def <- as.data.frame(out2)
def$weights <- as.factor(def$weights)
offset <- is.na(def$AVE_inner)
factorial_weights <- droplevels(def[!offset, ])
saveRDS(factorial_weights, file = "factorial_weights.RDS")
```

```{r}
lmWF0 <- lm(AVE_inner ~ 0+var3+var2+var1, data=factorial_weights)
lmWF <- lm(AVE_inner ~ 0+var3*var2*var1, data=factorial_weights)
anova(lmWF0, lmWF)
glance(lmWF)
tidy(lmWF) %>% arrange(desc(abs(estimate)))
```

```{r plot_factorial}
p <- ggplot(factorial_weights)
r1 <- p + geom_point(aes(cc1, AVE_inner, col = weights)) + guides(col = FALSE)
r2 <- p + geom_point(aes(cc1, AVE_outer, col = weights)) + guides(col = FALSE)
r3 <- p + geom_count(aes(AVE_inner, AVE_outer, col = weights))
q1 <- p + geom_point(aes(var1, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q2 <- p + geom_point(aes(var2, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q3 <- p + geom_point(aes(var3, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_outer, col = weights))
s1 <- p + geom_point(aes(var1, AVE_inner, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var1, AVE_inner, col = weights))
s2 <- p + geom_point(aes(var2, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_inner, col = weights))
s3 <- p + geom_point(aes(var3, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_inner, col = weights))
(r1 + r2 + r3) / (q1 + q2 + q3) / (s1 + s2 + s3)
```

```{r}
factorial_weights %>% 
  group_by(sign(vs12), sign(vs13), sign(vs23), weights) %>% add_count() %>% 
  filter(AVE_inner == max(AVE_inner)) %>% 
  arrange(as.numeric(weights), AVE_inner)
```

We can evaluate the influence of the scheme on the weights:
```{r}
def <- rbind(centroid_weights, factorial_weights, horst_weights)
def <- cbind(def, scheme= c(rep("centroid", nrow(centroid_weights)),
                            rep("factorial", nrow(factorial_weights)),
                            rep("horst", nrow(horst_weights))))
lmDef0 <- lm(AVE_inner~0+var1+var2+var3, data=def)
lmDef <- lm(AVE_inner~0+var1*var2*var3, data=def)
anova(lmDef0, lmDef)
glance(lmDef)
tidy(lmDef) %>% arrange(desc(abs(estimate)))
lmDef20 <- lm(AVE_inner~0+scheme+var1+var2+var3, data=def)
lmDef2 <- lm(AVE_inner~0+scheme*var1*var2*var3, data=def)
glance(lmDef2)
tidy(lmDef2) %>% arrange(desc(abs(estimate)))
```


## Conclusions

In general using only two weights leads to select models that account for more 
variance while, even if we can't see in here, the model having only 2 variables 
depends more on the weights regardless which scheme we use.
The weights that report more inner AVE are consistent for each scheme.
The distinction seems to be between adding a slighly relationship between CGH 
and GE or not. [CGE](https://en.wikipedia.org/wiki/Comparative_genomic_hybridization)
allows to find copy number variations, which might affect the gene expression, 
thus is reasonable to include in a model the interaction between copy number 
variation and gene expression, as well as it is unexpected that it might play a .
huge role But we can see that its role is not neglibile 

We can check them:
```{r}
C_model2 <- structure(c(0, 0, 1, 0, 0, 0.1, 1, 0.1, 0), .Dim = c(3L, 3L), .Dimnames = list(
    c("GE", "CGH", "y"), c("GE", "CGH", "y")))
C_model3 <- structure(c(0, 0.1, 1, 0.1, 0, 0.1, 1, 0.1, 0), .Dim = c(3L, 3L), .Dimnames = list(
    c("GE", "CGH", "y"), c("GE", "CGH", "y")))

centroid.2 <- sgcca(A, C_model2, c1 = shrinkage, ncomp = ncomp, 
      scheme = "centroid", verbose = FALSE)
centroid.3 <- sgcca(A, C_model3, c1 = shrinkage, ncomp = ncomp, 
      scheme = "centroid", verbose = FALSE)
plot(centroid.2$Y[[1]][, 1], centroid.2$Y[[2]][, 1], 
     col = Loc, pch = 16, xlab = "GE", ylab = "CGH", main = "Model 2")
plot(centroid.3$Y[[1]][, 1], centroid.3$Y[[2]][, 1], 
     col = Loc, pch = 16, xlab = "GE", ylab = "CGH", main = "Model 3")
```

Note that there are few differences between them, both of them allow to 
differentiate localization of the tumor by gene expression. Indeed, both of them
select the same genes `r length(intersect(names(comp.GE.3)[comp.GE.3 != 0], names(comp.GE.2)[comp.GE.2 != 0]))`,
but select different number of copy number variations (`r sum(comp.CGH.2 != 0)` and `r sum(comp.CGH.3 != 0)`, for model 2 and model 3), altough `r ength(intersect(names(comp.CGH.3)[comp.CGH.3 != 0], names(comp.CGH.2)[comp.CGH.2 != 0]))` are shared. 

## Session Info {-}

```{r session, echo=FALSE}
devtools::session_info()
```
