---
title: "Weight optimization with interaction"
author: "Lluís Revilla Sancho"
date: "`r date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, fig.width = 120, fig.height = 50, collapse = TRUE)
library("RGCCA")
library("gliomaData")
library("ggplot2")
library("patchwork")
library("integration")
theme_set(theme_bw())
```

# Introduction

In this vignette we explore the relationship between weights (and models) with the outcome of the integration using the glioma data available, with an interaction whitin the gene expression.

```{r introduction}
data(ge_cgh_locIGR, package = "gliomaData")
A <- ge_cgh_locIGR$multiblocks
Loc <- factor(ge_cgh_locIGR$y)
levels(Loc) <- colnames(ge_cgh_locIGR$multiblocks$y)
ncomp <- rep(1, length(A))
shrinkage <- c(.071, 1)
# rgcca algorithm using the dual formulation for X1 and X2 
# and the dual formulation for X3
A[[3]] = A[[3]][, -3]

# Prepare some preliminary design/weights
C <-  matrix(c(0, 0, 1, 0, 0, 1, 1, 1, 0), 3, 3)
dimnames(C) <- list(names(A), names(A))
```

# Weights

To test some weights we can use the RV index between the matrices

```{r RV, echo=FALSE}
D2 <- D <- C
for (i in seq_len(nrow(C))) {
  Ai <- scale(A[[i]], scale = FALSE)
  for (j in seq_len(i)) {
    Bj <- scale(A[[j]], scale = FALSE)
    if (i == j) {
      D2[i, i] <- 0
    } else {
      D2[i, j] <- MatrixCorrelation::RV(Ai, Bj)
      D2[j, i] <- D2[i, j]
    }
  }
}

```

The correlation between the first components of a multidimentional scaling plot:

```{r cor}
cmds <- lapply(A, function(x){cmdscale(dist(scale2(x)), k = 1)})
for (i in seq_len(nrow(C))) {
  for (j in seq_len(i)) {
    if (i == j) {
      D[i, i] <- 0
    } else {

      D[i, j] <- cor(cmds[[i]], cmds[[j]])
      D[j, i] <- D[i, j]
    }
  }
}
D <- abs(D)
```


Create the 1000 design matrix combinations

```{r}
nweight <- 11
weight <- seq(from = 0, to = 1, length.out = nweight)
C_list2 <- weight_design(11, 3)
C_list3 <- c(C_list2,
             lapply(C_list2, function(x){x[1, 1] <- weight[2];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[3];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[4];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[5];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[6];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[7];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[8];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[9];x}),
             lapply(C_list2, function(x){x[1, 1] <- weight[10];x}))
keep <- check_design(C_list3)
C_list3 <- C_list3[keep]
```


## The genes selected

To know if it really makes sense to change the matrix weights we can look for changes

```{r}
result.sgcca = sgcca(A, D, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
result.sgcca2 = sgcca(A, D2, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
result.sgcca3 = sgcca(A, C, c1 = shrinkage, ncomp = ncomp, 
                     scheme = "centroid", verbose = FALSE)
comp.weight <- function(x, y){
  xdiff <- x != 0
  ydiff <- y != 0
  intersect(names(x[xdiff]), names(y[ydiff]))
}

l <- list(result.sgcca, result.sgcca2, result.sgcca3)
outm <- matrix(0, ncol = length(l), nrow = length(l))
for (i in seq_len(length(l))) {
  for (j in seq_len(i)) {
    outm[i, j] <- length(comp.weight(l[[i]]$a[[1]][, 1], l[[j]]$a[[1]][, 1]))
    outm[j, i] <- outm[i, j]
  }
}
GE <- outm/max(outm)
dimnames(GE) <- list(c("D", "D2", "C"), c("D", "D2", "C"))
GE
```
We can see that with different weights different genes are selected, and if we look at the CGH:
```{r}
outm <- matrix(0, ncol = length(l), nrow = length(l))
for (i in seq_len(length(l))) {
  for (j in seq_len(i)) {
    outm[i, j] <- length(comp.weight(l[[i]]$a[[2]][, 1], l[[j]]$a[[2]][, 1]))
    outm[j, i] <- outm[i, j]
  }
}
CGH <- outm/max(outm)
dimnames(CGH) <- list(c("D", "D2", "C"), c("D", "D2", "C"))
CGH
```

We see they also change however almost all agree but there are some slight changes of up to 25%


```{r}
testing <- function(x, type) {
  result.sgcca <- RGCCA::sgcca(A, x, c1 = shrinkage, ncomp = c(1, 1, 1), 
                       scheme = type, verbose = FALSE)
  out <- analyze(result.sgcca)
  c(out, "var11" = result.sgcca$C[1, 1])
}

# Design matrix:
# 0.0 var12 var13
# var12 0.0  var23
# var13 var23  0.0
```

### Centroid scheme

Using the centroid scheme

```{r centroid}
keep <- vapply(C_list3, correct, logical(1L))
design <- C_list3[correct]
out <- sapply(design, testing, type = "centroid", USE.NAMES = FALSE)
out2 <- t(out)
def <- as.data.frame(out2)
def$weights <- as.factor(def$weights)
offset <- is.na(def$AVE_inner)
def <- droplevels(def[!offset, ])
saveRDS(def, file = "centroid_interactions.RDS")
```

We can explore by type of model how accurate is the model.
```{r}
library("dplyr")
def %>% 
  group_by(weights) %>% 
  add_count() %>% 
  dplyr::filter(AVE_inner == max(AVE_inner)) %>% 
  arrange(desc(AVE_inner), desc(as.numeric(weights))) %>% 
  select(-matches("vs"))
```
We can plot them

```{r plot_centroid}
p <- ggplot(def)
r1 <- p + geom_point(aes(cc1, AVE_inner, col = weights)) + guides(col = FALSE)
r2 <- p + geom_point(aes(cc1, AVE_outer, col = weights)) + guides(col = FALSE)
r3 <- p + geom_count(aes(AVE_inner, AVE_outer, col = weights))
p + geom_point(aes(AVE_inner, AVE_outer, col = weights)) + facet_wrap(~weights)
p+geom_point(aes(AVE_inner, cc1, col = weights))
q1 <- p + geom_point(aes(var1, AVE_outer, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q2 <- p + geom_point(aes(var2, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_outer, col = weights))
q3 <- p + geom_point(aes(var3, AVE_outer, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_outer, col = weights))
s1 <- p + geom_point(aes(var1, AVE_inner, col = weights))  + guides(col = FALSE) +
  geom_smooth(aes(var1, AVE_inner, col = weights))
s2 <- p + geom_point(aes(var2, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var2, AVE_inner, col = weights))
s3 <- p + geom_point(aes(var3, AVE_inner, col = weights)) + ylab("") + guides(col = FALSE) +
  geom_smooth(aes(var3, AVE_inner, col = weights))
(r1 + r2 + r3) / (q1 + q2 + q3) / (s1 + s2 + s3)
```

## Conclusions

In general using only two weights leads to select models that account for more 
variance while, even if we can't see in here, the model having only 2 variables 
depends more on the weights regardless which scheme we use.
The weights that report more inner AVE are consistent for each scheme.
The distinction seems to be between adding a slighly relationship between CGH 
and GE or not. [CGE](https://en.wikipedia.org/wiki/Comparative_genomic_hybridization)
allows to find copy number variations, which might affect the gene expression, 
thus is reasonable to include in a model the interaction between copy number 
variation and gene expression, as well as it is unexpected that it might play a .
huge role But we can see that its role is not neglibile 

We can check them:
```{r}
C_model2 <- matrix(0, ncol = 3L, nrow = 3L, dimnames = list(
    c("GE", "CGH", "y"), c("GE", "CGH", "y")))
m2 <- def[def$weights == "2", ]
m3 <- def[def$weights == "3", ]
C_model2 <- symm(C_model2, m2[which.max(m2$AVE_inner), grep("var", colnames(m2))])
C_model3 <- symm(C_model2, m3[which.max(m3$AVE_inner), grep("var", colnames(m3))])
centroid.2 <- sgcca(A, C_model2, c1 = shrinkage, ncomp = ncomp, 
      scheme = "centroid", verbose = FALSE)
centroid.3 <- sgcca(A, C_model3, c1 = shrinkage, ncomp = ncomp, 
      scheme = "centroid", verbose = FALSE)
plot(centroid.2$Y[[1]][, 1], centroid.2$Y[[2]][, 1], 
     col = Loc, pch = 16, xlab = "GE", ylab = "CGH", main = "Model 2")
plot(centroid.3$Y[[1]][, 1], centroid.3$Y[[2]][, 1], 
     col = Loc, pch = 16, xlab = "GE", ylab = "CGH", main = "Model 3")
```

Note that there are few differences between them, both of them allow to 
differentiate localization of the tumor by gene expression. Indeed, both of them
select the same genes `r length(intersect(names(comp.GE.3)[comp.GE.3 != 0], names(comp.GE.2)[comp.GE.2 != 0]))`,
but select different number of copy number variations (`r sum(comp.CGH.2 != 0)` and `r sum(comp.CGH.3 != 0)`, for model 2 and model 3), altough `r ength(intersect(names(comp.CGH.3)[comp.CGH.3 != 0], names(comp.CGH.2)[comp.CGH.2 != 0]))` are shared. 

# Bootstrapping the samples

```{r}

# Model 2
boot <- boot_sgcca(A, C_model2, shrinkage, 1000) # Bootstraps
saveRDS(boot, file = "bootstrap_model2_best.RDS") # Save
boot_evaluate(boot$STAB) # Plot some models

# Model 3
boot <- boot_sgcca(A, C_model3, shrinkage, 1000)
saveRDS(boot, file = "bootstrap_model3_best.RDS")
boot_evaluate(boot$STAB)
```

## Session Info {-}

```{r session, echo=FALSE}
devtools::session_info()
```
